---
title: "Homework 6"
author: Matthew T. Russell
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r, read in birthweight data and see if there are any NA values, message=FALSE}
birthweight_data <- read_csv("https://www.p8105.com/data/birthweight.csv")

birthweight_missing <- map_df(birthweight_data, ~sum(is.na(.))) %>% 
  pivot_longer(
    cols = babysex:wtgain,
    names_to = "variable", 
    values_to = "num_missing"
  )

birthweight_missing %>% kable()
```

From the table above, we can see that none of the variables contain any NA values. 

```{r, see if frace variable has any unknown, message=FALSE}
birthweight_data %>% 
  group_by(frace) %>% 
  summarize(
    total = n()
  ) %>% 
  kable()
```

Per the data description, we know that the variable correspond to the father's race has a value for observations where the value is unknown (`frace == 9`), However, from the table above, we can see that there aren't any observations that have this value. 

Now that we know there aren't any missing values, we can fit a model. 

```{r, tidying birthweight data, message=FALSE}
birthweight_df <-
  birthweight_data %>% 
  mutate(
    babysex = ifelse(babysex == 1, "male", "female") %>% 
      forcats::as_factor() %>% 
      relevel(ref = "male"),
    frace = case_when(
      frace == 1 ~ "white", 
      frace == 2 ~ "black", 
      frace == 3 ~ "asian", 
      frace == 4 ~ "puerto rican", 
      frace == 8 ~ "other"
    ) %>% 
      forcats::as_factor() %>% 
      relevel(ref = "other"),
    malform = ifelse(malform == 0, "absent", "present") %>% 
      forcats::as_factor() %>% 
      relevel(ref = "absent"), 
    mrace = case_when(
      mrace == 1 ~ "white", 
      mrace == 2 ~ "black", 
      mrace == 3 ~ "asian", 
      mrace == 4 ~ "puerto rican"
    ) %>% 
      forcats::as_factor() %>% 
      relevel(ref = "puerto rican")
  )
```

```{r, regression model, message=FALSE}
bwt_fit <- lm(bwt ~ fincome + frace + gaweeks + mrace + momage + smoken, data = birthweight_df)
```

